---
title: "Data Sources and Preparation"
output: 
  github_document:
    toc: yes
  html_document:
    toc: yes
version: 1.03
---

# SUMMARY
## I. Daily Data
#### i) [EDF](https://my.edfenergy.com/user/login)
Monthly csv downloads of daily household gas and electricity consumption data were aggregated into a spreadsheet and exported to a csv file. The data covers the period 15/03/2021 - 30/6/2023. Pivot tables were used to cross check monthly totals with the EDF energy hub to highlight any copy paste errors.

Changelog:  
31/03/2023 gas usage updated 29/5/23 from previous estimate of 30 to 27.63 Kwh  
04/06/2023 2023 April and May figures added  
10/07/2023 June figures added  

#### ii) Solis app
Solar PV was installed 20/9/2022. Daily yield data was manually transcribed from the Solis app to a spreadsheet. Monthly totals were checked against the app to reveal transcription errors. This highlighted some discrepancies between the monthly and aggregated daily totals within the app. The assumption was made that the monthly total was affected less by wi-fi connectivity issues and therefore recorded more accurately. In consequence some "zero yield" days were filled with estimates to match the monthly total. 
It was also noticed that the most recent days of yield data were later updated to slightly higher figures. The reason was not clear. Possibly some type of calibration or reconciliation with meter readings.
The yield data was combined with degree days and supplier consumption data into one dataset of daily records.

Changelog:  
Updates from "zero yield" days to bring monthly total in line with app:-  
22/11/22 yield estimate = 3  
15/2/23 yield estimate = 5  
16/2/23 yield estimate = 5  
17/2/23 yield estimate = 5  
18/2/23 yield estimate = 4.3  

10/7/23 June 2023 figures added
  
#### iii) [Degree days](https://www.degreedays.net/)
Degree days were calculated using www.degreedays.net with a base temperature of 15.5 degrees C. Temperature data was taken from Liverpool airport weather station. Data was downloaded in csv format and added into the spreadsheet of daily data.

Changelog:  
10/7/23 June 2023 figures added

## II. Hourly Data

#### [N3rgy](https://data.n3rgy.com/consumer/home)
Personal half hourly energy consumption data were downloaded in 2-monthly batches. (the website had a 90 days download limit). The first available data was October 2021. The data were aggregated into one dataset of hourly data in local time using R (see below) and written to a csv. 

Checks were made for duplicates and missing entries, and a few gaps were filled. 
Cross checks for the electric data were made against the supplier provided daily data. This highlighted a few discrepancies when EDF had supplied an estimate rather than actual meter read. The largest of these was 2.2 Kwh on 20/10/2021.  Rounding errors may also contribute to smaller discrepancies; The n3rgy data is provided to 3dp whereas the EDF daily data is 2dp.  
Monthly totals were all within 1% except for March 2023 (electric, n3rgy was missing 1.5 days data), and May, June 2023 for gas (out by under 2.5 kwh for each). The total difference over the 20 months of available n3rgy data was under 20KWh for electric. This is less than half a percent so not expected to impact this study.
Samples of hourly figures were also compared with EDF's energy hub data and are largely consistent with hourly discrepancies less than .01 KWh.

The half hourly data was aggregated by hour in local time.

Changelog:  
  10 missing records created for:-  
  "2021-11-12 07:30:00 UTC" "2022-01-10 03:00:00 UTC" "2022-03-04 06:00:00 UTC"
  "2022-05-06 08:00:00 UTC" "2022-07-14 08:00:00 UTC" "2022-09-30 09:30:00 UTC"
  "2022-12-01 05:30:00 UTC" "2023-01-19 19:00:00 UTC" "2023-03-09 00:00:00 UTC"
  "2023-04-29 06:30:00 UTC"  
  The electric consumption entry was filled by repeating the previous record.  

  4/6/23 Hourly summary amended from hour ending to the centre of the range to align with supplier data.  
  11/7/23 June 2023 figures added  
  13/7/23 gas data added from 10/11/22  NB measured in cubic metres so is converted into Kwh.This may introduce some small discrepancies with supplier data as calorific values vary.   
  20/8/23 checks added to demonstrate correct aggregation by hour
  
## III. Monthly Data
#### [Personal solar yield estimate](https://re.jrc.ec.europa.eu/pvg_tools/en/)
Location, size of PV array, azimuth and inclination were used to personalise solar yield estimates from the PVGIS tool from the European Commission. (3.95 KWp, 14% system loss, slope 30, azimuth -35). PVGIS uses satellite solar radiation data with climate reanalysis models to produce energy yield information. Monthly data were downloaded from the website as a double tab delimited file. This was imported and prepared in R by selecting and renaming columns. 
The annual estimate of generation is 3422 KWh with standard deviation given as 117 Kwh over the 15 years of the solar radiation database.

---


```{r message=FALSE, warning=FALSE}
library(tidyverse) # core 
library(skimr)
library(data.table)
```

# 1. Load prepared daily data

```{r}
daily_data <- read_csv("Data/230630_daily.csv", col_types = cols(date = col_date(format = "%d/%m/%y")))

# quick overview
skim_without_charts(daily_data)

# check for missing entries
diffs <- daily_data$date - shift(daily_data$date)

daily_data[c(which(diffs>1)-1, which(diffs>1)),] %>%
  arrange(date)
```
No missing dates

```{r}
tail(daily_data)
```

# 2. Load and prepare half hourly (n3rgy) data

```{r}
# Half hourly electric import data in 2 month batches
dat1 <- read_csv("Data/n3rgy/20211001-20211201.csv", show_col_types = FALSE)
dat2 <- read_csv("Data/n3rgy/20211201-20220201.csv", show_col_types = FALSE)
dat3 <- read_csv("Data/n3rgy/20220201-20220401.csv", show_col_types = FALSE)
dat4 <- read_csv("Data/n3rgy/20220401-20220601.csv", show_col_types = FALSE)
dat5 <- read_csv("Data/n3rgy/20220601-20220801.csv", show_col_types = FALSE)
dat6 <- read_csv("Data/n3rgy/20220801-20221001.csv", show_col_types = FALSE)
dat7 <- read_csv("Data/n3rgy/20221001-20221201.csv", show_col_types = FALSE)
dat8 <- read_csv("Data/n3rgy/20221201-20230201.csv", show_col_types = FALSE)
dat9 <- read_csv("Data/n3rgy/20230201-20230401.csv", show_col_types = FALSE)
dat10 <- read_csv("Data/n3rgy/20230401-20230601.csv", show_col_types = FALSE)
dat11 <- read_csv("Data/n3rgy/20230601-20230701.csv", show_col_types = FALSE)

# Combine into one dataframe and sort by date
elec_halfhr <- rbind(dat1, dat2, dat3, dat4, dat5, dat6, dat7, dat8, dat9, dat10, dat11) %>% 
  select('timestamp (UTC)' ,'energyConsumption (kWh)' ) %>% 
  rename( 'UTC_time' = 'timestamp (UTC)') %>%
  rename( 'elec_import_kwh' = 'energyConsumption (kWh)') %>%
  arrange(UTC_time)
```

```{r}
# Half hourly gas import data
datg1 <- read_csv("Data/n3rgy/gas/20221110-20221201.csv", show_col_types = FALSE)
datg2 <- read_csv("Data/n3rgy/gas/20221201-20230201.csv", show_col_types = FALSE)
datg3 <- read_csv("Data/n3rgy/gas/20230201-20230401.csv", show_col_types = FALSE)
datg4 <- read_csv("Data/n3rgy/gas/20230401-20230601.csv", show_col_types = FALSE)
datg5 <- read_csv("Data/n3rgy/gas/20230601-20230701.csv", show_col_types = FALSE)

gas_halfhr <- rbind(datg1, datg2, datg3, datg4, datg5) %>% 
  select('timestamp (UTC)' ,'energyConsumption (m3)' ) %>% 
  rename( 'UTC_time' = 'timestamp (UTC)') %>%
  rename( 'gas_m3' = 'energyConsumption (m3)') %>%
  arrange(UTC_time)

## convert m3 to kwh 
# (https://www.gov.uk/guidance/gas-meter-readings-and-bill-calculation)
gas_halfhr$gas_kwh <- gas_halfhr$gas_m3*39.5*1.02264/3.6

```

## Checks
```{r}
skim_without_charts(elec_halfhr)
str(elec_halfhr)
tail(elec_halfhr)

skim_without_charts(gas_halfhr)
str(gas_halfhr)
tail(gas_halfhr)

# tidy up
rm(dat1,dat2,dat3,dat4,dat5,dat6,dat7,dat8,dat9, dat10, dat11)
rm(datg1, datg2, datg3, datg4, datg5)
```


```{r}
# check for missing entries
diffs = elec_halfhr$UTC_time - shift(elec_halfhr$UTC_time)
elec_halfhr[c(which(diffs>30)-1, which(diffs>30)),] %>%
  arrange(UTC_time)

diffs = gas_halfhr$UTC_time - shift(gas_halfhr$UTC_time)
gas_halfhr[c(which(diffs>30)-1, which(diffs>30)),] %>%
  arrange(UTC_time)
```

It turns out there are 10 isolated missing electric measurements and 1 missing gas measurement up until 1/7/23. There is no obvious reason - perhaps due to lost connectivity. 
All electric records are missing between 21/3/23 13:00 and 23/3/23 00:00.
Gas records are only available from 10/11/22.

### Deal with missing data

Create the single missing measurements by repeating the previous value:
```{r}
fill_dates <- c(ymd_hm("2021-11-12 07:30"), ymd_hm("2022-01-10 03:00"), ymd_hm("2022-03-04 06:00"), ymd_hm("2022-05-06 08:00"), ymd_hm("2022-07-14 08:00"), ymd_hm("2022-09-30 09:30"), ymd_hm("2022-12-01 05:30"), ymd_hm("2023-01-19 19:00"), ymd_hm("2023-03-09 00:00"), ymd_hm("2023-04-29 06:30"))

filled_elec_halfhr <- rbind(
    elec_halfhr, 
    data.frame(
      UTC_time = c(ymd_hm("2021-11-12 07:30"), ymd_hm("2022-01-10 03:00"), ymd_hm("2022-03-04 06:00"), ymd_hm("2022-05-06 08:00"), ymd_hm("2022-07-14 08:00"), ymd_hm("2022-09-30 09:30"), ymd_hm("2022-12-01 05:30"), ymd_hm("2023-01-19 19:00"), ymd_hm("2023-03-09 00:00"), ymd_hm("2023-04-29 06:30")), 
      elec_import_kwh=NA)) %>%
  arrange(UTC_time)

filled_gas_halfhr <- rbind(
  gas_halfhr,
  data.frame(
    UTC_time = c(ymd_hm("2023-05-30 13:00")), 
    gas_m3=NA, gas_kwh=NA)) %>%
  arrange(UTC_time)


# Fill missing data by repeating previous value.
ixs <- which(is.na(filled_elec_halfhr$elec_import_kwh))
filled_elec_halfhr$elec_import_kwh <- 
  nafill(filled_elec_halfhr$elec_import_kwh, "locf")

filled_gas_halfhr$gas_kwh <- 
  nafill(filled_gas_halfhr$gas_kwh, "locf")

# check the records filled
filled_elec_halfhr[ixs,] 
```
10 records created.
```{r}
# check for duplicates
nrow(filled_elec_halfhr[duplicated(filled_elec_halfhr$UTC_time), ])
nrow(filled_gas_halfhr[duplicated(filled_gas_halfhr$UTC_time), ])
```

```{r}
# check again for missing entries
diffs = filled_elec_halfhr$UTC_time - shift(filled_elec_halfhr$UTC_time)

filled_elec_halfhr[c(which(diffs>30)-1, which(diffs>30)),] %>%
  arrange(UTC_time)

diffs = filled_gas_halfhr$UTC_time - shift(filled_gas_halfhr$UTC_time)

filled_gas_halfhr[c(which(diffs>30)-1, which(diffs>30)),] %>%
  arrange(UTC_time)
```

All looks good. Just the large gap between 21st and 23rd March 2023 in the electric data.

```{r}
skim_without_charts(filled_elec_halfhr)
str(filled_elec_halfhr)

str(filled_gas_halfhr)
```
```{r}
# combine gas and elec data
n3rgy <- merge(filled_elec_halfhr, filled_gas_halfhr,
                    by="UTC_time",all=TRUE)

# fill NAs following merge with 0 (21,22/3/23 for elec) 
n3rgy$elec_import_kwh <- 
  nafill(n3rgy$elec_import_kwh, fill=0)
```

### Look for discrepancies with supplier data

Compare daily totals with dataset from EDF
```{r}
# n3rgy data
n3rgy$date <- as.Date(n3rgy$UTC_time)
n3rgy_summary <-
  n3rgy %>%
  group_by(date) %>%
  summarise(n3rgy_elec_import_kwh=sum(elec_import_kwh),
            n3rgy_gas_kwh=sum(gas_kwh),
            )

daily_summary <- merge(n3rgy_summary, daily_data, by="date")


# looking for discrepancies 

# days where the EDF and n3rgy data disagree by more than 1kwh
daily_summary[which(abs(daily_summary$n3rgy_elec_import_kwh - daily_summary$elec_usage_kwh) > 1), c("date","n3rgy_elec_import_kwh", "elec_usage_kwh")]

daily_summary[which(abs(daily_summary$n3rgy_gas_kwh - daily_summary$gas_usage_kwh) > 1), c("date","n3rgy_gas_kwh", "gas_usage_kwh")]
```

The days where supplier and n3rgy data disagree most for electricity consumption coincide with days when the supplier provided estimates and where the n3rgy dataset was missing records. This does not seem to be the case for gas data.Possibly the conversion

Compare monthly totals with daily dataset from EDF
```{r warning=FALSE}
# EDF data
daily_data$yr_mth <- paste(year(daily_data$date), month(daily_data$date), sep='-')
EDF_summary <-
  daily_data %>%
  group_by(yr_mth) %>%
  summarise(EDF_elec_import_kwh=sum(elec_usage_kwh),
            EDF_gas_import_kwh=sum(gas_usage_kwh),
           # EDF_count=n() 
            )

# n3rgy data
n3rgy$yr_mth <- paste(year(n3rgy$UTC_time),
                      month(n3rgy$UTC_time), sep='-')
n3rgy_summary <-
  n3rgy %>%
  group_by(yr_mth) %>%
  summarise(n3rgy_elec_import_kwh=sum(elec_import_kwh),
            n3rgy_gas_kwh=sum(gas_kwh),
          #  n3rgy_count=n() 
            )


monthly_summary <- merge(n3rgy_summary, 
             EDF_summary, by="yr_mth")
monthly_summary$elec_diff <- monthly_summary$EDF_elec_import_kwh -
  monthly_summary$n3rgy_elec_import_kwh

monthly_summary$gas_diff <- monthly_summary$EDF_gas_import_kwh -
  monthly_summary$n3rgy_gas_kwh

# looking for discrepancies > 1 %
monthly_summary[which(abs(monthly_summary$elec_diff/monthly_summary$EDF_elec_import_kwh) > 0.01), c("yr_mth","elec_diff","EDF_elec_import_kwh")]
monthly_summary[which(abs(monthly_summary$gas_diff/monthly_summary$EDF_gas_import_kwh) > 0.01),c("yr_mth","gas_diff","EDF_gas_import_kwh")]
```



The EDF and n3rgy data are not identical but monthly totals agree to within a few KWh - less than 1% on electric for all months bar March 2023 which is missing 1.5 day's data in the n3rgy dataset.

Total difference (Kwh) between supplier and n3rgy electric data (Oct 2021 - June 2023):
```{r}
# Total difference between supplier and n3rgy electric data 
round(sum(daily_summary$n3rgy_elec_import_kwh) - sum(daily_summary$elec_usage_kwh),2)

```
## Aggregate by hour in local time

```{r}
# convert to local time and aggregate by hour
n3rgy_hourly <- n3rgy %>%
   mutate(local_time=format(as.POSIXct(with_tz(n3rgy$UTC_time, "Europe/London") ), "%Y-%m-%d %H")) %>%
  # aggregate st 01:00 + 01:30 => 02:00 if BST else => 01:00 (centre of range assuming data is hour ending)
  group_by(local_time) %>%
  summarise(elec_import_kwh=sum(elec_import_kwh),
            gas_kwh=round(sum(gas_kwh),3)) %>%
  mutate(local_time=as.POSIXct(local_time, tz="Europe/London", format = "%Y-%m-%d %H")) 
```

Doing a check to be sure the aggregation is as expected and timezones are dealt with correctly..
```{r}
# Check aggregation as expected and timezones dealt with correctly
# (NB normally not a good idea to test floats for equality
# Here testing result is expected sum)

# BST ended 31/10/21 and 30/10/22 => 
# TEST BST: 31/10/21 00:00 local time (BST) is sum of 30/10/21 23:00 and 23:30 in UTC
n3rgy_hourly$elec_import_kwh[which(
  n3rgy_hourly$local_time == as.POSIXct("2021-10-31 00:00:00", tz="Europe/London")
  )] == 
n3rgy$elec_import_kwh[which(n3rgy$UTC_time == as.POSIXct("2021-10-30 23:00:00", tz="UCT"))] + n3rgy$elec_import_kwh[which(n3rgy$UTC_time == as.POSIXct("2021-10-30 23:30:00", tz="UCT"))]

# TEST Autumn clock change: 31/10/21 01:00 local time is sum of 00:00-01:30 in UTC,
n3rgy_hourly$elec_import_kwh[which(
  n3rgy_hourly$local_time == as.POSIXct("2021-10-31 01:00:00", tz="Europe/London")
  )] == 
sum( n3rgy$elec_import_kwh[which(
  n3rgy$UTC_time >= as.POSIXct("2021-10-31 00:00:00", tz="UCT") & 
  n3rgy$UTC_time <= as.POSIXct("2021-10-31 01:30:00", tz="UCT")
  )] 
)

# TEST GMT: 31/10/21 02:00 local time (after clock put back to GMT) is 
# sum of 02:00 and 02:30 UTC
n3rgy_hourly$elec_import_kwh[which(
  n3rgy_hourly$local_time == as.POSIXct("2021-10-31 02:00:00", tz="Europe/London")
  )] ==  
n3rgy$elec_import_kwh[which(n3rgy$UTC_time == as.POSIXct("2021-10-31 02:00:00", tz="UCT"))] + n3rgy$elec_import_kwh[which(n3rgy$UTC_time == as.POSIXct("2021-10-31 02:30:00", tz="UCT"))]

## show the records for a visual check
n3rgy_hourly[which(
  n3rgy_hourly$local_time >= as.POSIXct("2021-10-30 23:00:00", tz="UCT") &
  n3rgy_hourly$local_time < as.POSIXct("2021-10-31 03:00:00", tz="UCT")),]

n3rgy[which(
  n3rgy$UTC_time >= as.POSIXct("2021-10-30 23:00:00", tz="UCT") & 
  n3rgy$UTC_time < as.POSIXct("2021-10-31 03:00:00", tz="UCT")),]

# BST started 27/3/22, 26/3/22 =>
# there will be no record for 2022-03-27 01:00:00 in local time
# NB when requested directly R appears to repeat the record for 00:00, even tho the record is not in the table!
# TEST Spring clock change: Sum of 27/3/22 00:00 and 01:00 in local time is same as 00:00 
sum(n3rgy_hourly$elec_import_kwh[which(
  n3rgy_hourly$local_time >= as.POSIXct("2022-03-27 00:00:00", tz="Europe/London") & 
  n3rgy_hourly$local_time <= as.POSIXct("2022-03-27 01:00:00", tz="Europe/London")
  )] 
) == n3rgy_hourly$elec_import_kwh[which(
  n3rgy_hourly$local_time == as.POSIXct("2022-03-27 00:00:00", tz="Europe/London")
  )] 

sum(n3rgy_hourly$gas_kwh[which(
    n3rgy_hourly$local_time >= as.POSIXct("2023-03-26 00:00:00", tz="Europe/London") & 
    n3rgy_hourly$local_time <= as.POSIXct("2023-03-26 01:00:00", tz="Europe/London")
  )] 
) == 
  n3rgy_hourly$gas_kwh[which(
  n3rgy_hourly$local_time == as.POSIXct("2023-03-26 00:00:00", tz="Europe/London")
  )] 

## show the records for a visual check
n3rgy_hourly[which(
  n3rgy_hourly$local_time >= as.POSIXct("2023-03-26 00:00:00", tz="UCT") & 
  n3rgy_hourly$local_time < as.POSIXct("2023-03-26 03:00:00", tz="UCT")),]

n3rgy[which(
  n3rgy$UTC_time >= as.POSIXct("2023-03-26 00:00:00", tz="UCT") & 
  n3rgy$UTC_time < as.POSIXct("2023-03-26 03:00:00", tz="UCT")),]

```

Looks OK so write to csv and tidy up
```{r}
write.csv(n3rgy_hourly, file="Data/230630_hourly.csv", row.names=FALSE)

# tidy up some temporary files
rm(elec_halfhr, gas_halfhr, filled_elec_halfhr, filled_gas_halfhr, EDF_summary, n3rgy_summary, diffs, ixs)
```

# 3. Load and prepare solar yield estimate data

Take a quick visual check of loaded data
```{r}
solar_estimate_pvgis <- read.csv(
  "Data/solar estimate/PVdata_SA2_crystSi_3kWp_14_30deg_-35deg.csv",
                                 skip=9, nrows=12, sep='\t')

solar_estimate <- solar_estimate_pvgis %>% 
  select('Month' ,'E_d', 'E_m','SD_m' ) %>% 
  rename( 'month' = 'Month') %>%
  rename( 'avg_daily_yield_kwh' = 'E_d') %>%
  rename( 'avg_monthly_yield_kwh' = 'E_m') %>%
  rename( 'sd_monthly_yield_kwh' = 'SD_m') %>%
  arrange('month')


solar_estimate
```

```{r}
#write.csv(solar_estimate, file="Data/PVGIS_solar_estimate_mthly.csv", row.names=FALSE)
```
